---
title: "Final Project - Group 3"
author: "Anthony Vallejo & Vanessa Johnson "
date: "June 24, 2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(ggplot2)
library(scales)

```

Will read the NYC Data that was provided to use

```{r load-data}
nyc_office_data <- read_csv("CCRB%20Complaint%20Database%20Raw%2004.20.2021.csv")
```

Gets the New York Data and creates its respective plot that was illustrated in the published article

``` {r get the New York City Decile Plot}
df1 <- nyc_office_data %>% group_by(ComplaintID,OfficerID) %>% summarize(Complaints = n()) 

df2 <- df1 %>% group_by(OfficerID) %>% summarize(Complaints_number = n()) %>% arrange(Complaints_number) %>% 
  mutate(complain_sum = cumsum(Complaints_number)/sum(Complaints_number)) %>% 
  summarize(cum_sum = quantile(complain_sum, probs = seq(0,1, 0.1)))

percent_of_complaints <- data.frame(diff(as.matrix(df2))) %>% mutate(rank = row_number()) 

ggplot(percent_of_complaints, aes(x = rank, y = cum_sum, fill = cum_sum)) +
  geom_bar(stat = "identity") + scale_x_continuous(breaks = scales::pretty_breaks(n = 10), name = "Ranks in Deciles" ) +
  scale_y_continuous(name = 'Percent of Complaints', labels = percent) + ggtitle("Distribution of Civilian Misconduct Complaints") +
  theme(axis.text.x =  element_text(angle = 90))
```

Allegations are grouped into four categories (FADO). Our goal was to illustrate how these allegations
have shifted since 2006. 

``` {r graphs the types of Allegations over time}

df3 <- nyc_office_data %>% group_by(FADOType) %>% summarize(QTY = n()) %>% arrange(QTY) 
#%>% ggplot(aes(x = FADOType, y = QTY, fill = FADOType)) + geom_bar(stat = "identity")

df4 <- nyc_office_data %>% 
  mutate(year = as.numeric(format(parse_date(IncidentDate, format = "%m/%d/%Y"),"%Y"))) %>% 
  group_by(year, FADOType) %>% summarize(Occurances = n()) %>% arrange(year) %>% filter(year >= 2006) 

ggplot(df4, aes(x = year, y = Occurances, color = FADOType)) + geom_line() +
  scale_x_continuous(breaks = seq(2006, 2021, by = 1)) + theme(axis.text.x =  element_text(angle = 90)) + geom_point() + scale_y_continuous(name = 'Occurances of Complaints') + ggtitle("Police Complaints Since 2006")

```

For our narrow question, we wanted to visualize how force allegations related with the deciles of complaints that we computed in our first graph. They turned out to be oddly similar - leading us to infer that police officers with a greater amount of complaints also offended with greater physical force 

``` {r graphs the Force Allegations over time}
df5 <- nyc_office_data %>% filter(FADOType == "Force") %>% group_by(OfficerID, ComplaintID) %>%
  summarize(num_complaints = n()) %>% group_by(OfficerID) %>% 
  summarize(com_per_officer = n()) %>% arrange(com_per_officer) %>%
  mutate(force_complain_sum = cumsum(com_per_officer)/sum(com_per_officer))%>%
  summarize(force_cum_sum = quantile(force_complain_sum, probs = seq(0,1, 0.1)))

percent_of_force_complaints <- data.frame(diff(as.matrix(df5))) %>% mutate(rank = row_number()) 

ggplot(percent_of_force_complaints, aes(x = rank, y = force_cum_sum, fill = force_cum_sum)) +
  geom_bar(stat = "identity") + scale_x_continuous(breaks = scales::pretty_breaks(n = 10), name = "Ranks in Deciles" ) +
  scale_y_continuous(name = 'Percent of Force Complaints', labels = percent) + ggtitle("Distribution of Civilian Misconduct Force Complaints") +
  theme(axis.text.x =  element_text(angle = 90))
``` 
